---
title: "PRIME - Papillomavirus Rapid Interface for Modelling and Economics"
author: "LSHTM"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRIME Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# load package - prime2
library(prime2)
```

## PRIME
PRIME stands for “Papillomavirus Rapid Interface for Modelling and Economics”. 
The R package (prime2) is based of the spreadsheet-based tool (see http://primetool.org). 

PRIME provides users with an idea of: 

* The magnitude of the burden of cervical cancer
* The impact of introducing HPV vaccination for girls prior to sexual debut
* Healthcare costs incurred as a result of cervical cancer treatment
* Costs associated with vaccination
* Long-term savings which may result from a vaccination program


### Functions
The main PRIME functions are:

* RunCohort
* RunCountry
* analyseCosts

Other functions include:

* writelog
* RegisterBatchData
* RegisterBatchDataGavi
* BatchRun
* OutputGavi
* getISO3
* checkSize
* lifeTable
* ageCoverage
* propSexDebut
* dtColMatch
* dtAggregate
* monetary_to_number

R code for the functions are in the R-folder of the package, i.e. ./R/

Documentation of a function can be viewed, by running ?functionname in the console, i.e. ?RunCohort
```{r eval=FALSE}

# View documentation of RunCohort function
?RunCohort

```

The source code of a function can be viewed by running the functionname without any brackets in the console.
```{r eval=FALSE}

# View source code of RunCohort function
RunCohort

```


### Data
Several data sets are included in the package. The data files (.rda files) are stored as data tables (data.table) in the data folder of the package, i.e. ./data/

* data.cecx_1y_prevalence
* data.cecx_3y_prevalence
* data.cecx_5y_prevalence
* data.costcecx
* data.countryname
* data.global
* data.incidence
* data.mortall
* data.mortall.unwpp.mx
* data.mortcecx
* data.pop
* data.popproj
* data.quality
* data.sexual_debut
* data.valid

Documentation of a data set can be viewed, by running ?dataset-name in the console, i.e. ?data.incidence
```{r eval=FALSE}

# View documentation of data.incidence data set
?data.incidence

```

A data set can be viewed by typing the data file name in the console.
```{r eval=FALSE}

# View data set for age-specific incidence of cervical cancer in different countries (estimated by Globocan)
data.incidence

```


### Examples
Use RunCountry to run PRIME for a single birthcohort in a single country. 

* Countries are referenced by their ISO3 code.
```{r}
RunCountry ("RWA")
```

* This returns the age-specific rates for a country, given the values passed to RunCountry. 
* The size of the cohort at each age is given in the cohort_size column. 
* In order to get the number of people rather than the rates, the columns can be multiplied with the cohort_size.
```{r}
result <- RunCountry ("RWA")
result [,c("vaccinated", "immunized", "inc.cecx", "mort.cecx", "lifey","disability", "cost.cecx")] <- result[, cohort_size] * result[,c("vaccinated", "immunized", "inc.cecx", "mort.cecx", "lifey", "disability", "cost.cecx")]
result [age>15]
```

If you don't know the exact ISO3 code for a country, you can try to look it up using the getISO3 function.
```{r}
getISO3 ("Malawi")
```

If you want to analyse the costs/benefits of vaccinating this specific cohort, set analyseCosts to TRUE.
```{r}
RunCountry ("MWI", analyseCosts=TRUE)
```


RunCountry automatically assumes 100% coverage, 100% vaccine-efficacy, and age of vaccination at 12.

* If no year is provided, it will assume that vaccination will occur this year. Year can be provided as year-of-vaccination or birthcohort-year (i.e. year in which girls of this birth-cohort was born).
* Tweak the different values by passing them to the function. 
* See ?RunCountry for all options.
```{r}
RunCountry("MWI", analyseCosts=TRUE, discounting=TRUE, agevac=10, agecohort=10, vaceff_beforesexdebut=0.98, vaceff_aftersexdebut=0, cov=0.75, year_vac=2020)
```

* In essence, RunCountry() calls RunCohort(), and will provide it with country-specific estimates.
* When analyseCosts is set to TRUE, RunCountry() calls analyseCosts() and passes the results of RunCohort(). 
* You could also run it yourself, providing your own estimates for vaccine-costs and GDP per capita.

```{r}
analyseCosts (
  RunCountry("MWI", analyseCosts=FALSE, discounting=TRUE, agevac=10, agecohort=10, vaceff_beforesexdebut=0.98, vaceff_aftersexdebut=0, cov=0.75, year_vac=2020),
  vaccine_cost=100,
  gdp_per_capita=561
)
```

* Lifetables are used to follow the size of a birth-cohort over time. 
* They are generated based on values for the qx-column (i.e. probability of dying by age). 
* The lifeTable() function makes it easy to generate a lifetable.
```{r}
qx <- unlist(data.mortall[iso3=="MWI", as.character(0:100), with=F], use.names=F)
lifeTable(qx)
```

* RunCohort() uses this lifetable. It calculates age-specific coverage-rates using the ageCoverage() function.
```{r}
ages <- c(0:100)
routine_coverage <- 0.75
vaccine_efficacy_nosexdebut <- 0.99
vaccine_efficacy_sexdebut <- 0
lifetab <- lifeTable(unlist(data.mortall[iso3=="MWI", as.character(0:100), with=F], use.names=F), 9)
cohort <- unlist(data.popproj[iso3=="MWI", "2020"], use.names=F)
agevac <- 9
ageCoverage(ages, routine_coverage, vaccine_efficacy_nosexdebut=1, vaccine_efficacy_sexdebut=0, -1, lifetab, cohort, agevac, country_iso3="MWI")
```

* Data can be aggregated using the dtAggregate() function.
```{r}
result <- RunCountry ("MWI", analyseCosts=FALSE, discounting=TRUE, agevac=10, agecohort=10, vaceff_beforesexdebut=0.98, vaceff_aftersexdebut=0, cov=0.75, year_vac=2020)

result [,c("vaccinated", "immunized", "inc.cecx", "mort.cecx", "lifey", "disability", "cost.cecx")] <- result [, cohort_size] * result[,c("vaccinated", "immunized", "inc.cecx", "mort.cecx", "lifey", "disability", "cost.cecx")]

# aggregate results on age
result <- dtAggregate (result, "age", id.vars=c("scenario", "type") )
result
```


